Data center infrastructure is among the fastest-growing sites of industrial capital formation in OECD economies — absorbing an estimated \$930 billion in hyperscaler capex over six years, consuming 400~TWh of electricity annually, and generating sustained community contestation from Loudoun County, Virginia to the outskirts of Dublin. It is, in other words, precisely the kind of socio-material transition that innovation and sustainability transitions research was built to explain. Yet a bibliometric audit of 4,757 publications on data centers (OpenAlex, 2010--2025) finds zero engagement with the Multi-Level Perspective, Technological Innovation Systems, Strategic Niche Management, or sociotechnical regimes — the standard tools for analysing infrastructure transitions of comparable scale.

This paper addresses the gap. We propose \textit{datacentering} — a gerund foregrounding the active, contested process of assembling digital infrastructure — as a conceptual entry point, and develop a multimodal research agenda structured around the three dimensions of complex adaptive systems (CAS): how actors construct imaginaries that render infrastructure legible or illegible (WP1--2), how ownership hierarchies shape governance capacity (WP3), and how contestation signals produce or fail to produce regime response (WP4). Seed results across four jurisdictions illustrate the analytical purchase: an Environmental Claim Index divergence of $+165.7$ for the largest US hyperscaler suggests systematic decoupling of corporate sustainability claims from territorial accountability; an ownership network of 20 nodes and 35 edges shows how REIT structures dissolve governance counterparties; and a contestation corpus of 58 events demonstrates that Amsterdam's moratorium followed signal onset by under a year while Northern Virginia, with comparable signal intensity, produced no regime response — a difference explained by ownership structure, not community mobilisation.

We conclude that the CAS-like character of these assemblages demands collective, open knowledge production. The paper is accompanied by a community-maintained facility index, shared analytical pipelines, and open data, on the grounds that no single research team can build this cartography alone.
