%% Section 1 — Introduction
\section{Introduction}
\label{sec:introduction}

The infrastructure running the digital economy occupies a peculiar position in the scholarly literature: it is studied intensively in engineering and computer science, acknowledged briefly in environmental economics, contested occasionally in urban planning, but almost entirely absent from the fields — sustainability transitions, innovation systems, science and technology studies — that have developed the most sophisticated tools for understanding socio-material transformation. This paper begins from the premise that this absence is a problem for the field, not merely a gap for a research agenda to fill, and that correcting it requires both a conceptual reorientation and a methodological apparatus capable of reaching across the opacity that operators have deliberately constructed around their infrastructure.

Data centers now consume an estimated 400~TWh of electricity per year globally \citep{iea2024electricity}, a figure that makes them, collectively, among the fastest-growing industrial electricity users in OECD economies. They are physical: vast server halls, cooling towers that draw millions of litres of water daily, diesel generators staged for grid interruption, land parcels the size of small towns assembled quietly through option agreements and shell companies. They are contested: planning disputes, noise ordinances, water rights challenges, and community campaigns are a feature of markets from Loudoun County, Virginia to the outskirts of Dublin. They are governed — or systematically not governed — through structures that range from Real Estate Investment Trust (REIT) ownership dispersed across pension funds and retail investors to municipal moratoriums negotiated with named operators. And they are opaque by design: the term ``the cloud'' is not merely a marketing metaphor but an ideological instrument that naturalises infrastructure, makes it seem weightless and placeless, and thereby insulates it from the territorial accountability that its material footprint would otherwise attract.

\input{sections/box_megaprojects}

A bibliometric audit of 4,757 publications with ``data center'' in the title on OpenAlex (2010--2025) makes the disciplinary gap visible in quantitative terms. The literature has grown by 372\% over fifteen years and organised itself into three structurally disconnected silos: technical efficiency (approximately 279 papers focused on cooling, power management, and hardware optimisation), environmental impact (approximately 114 papers, primarily on carbon footprint and energy sourcing), and socio-political dynamics (approximately 15 papers, including the rich ethnographic literature of \citet{vonderau2019storing} and \citet{holt2015vonderau}). Zero papers in this corpus engage the Multi-Level Perspective, Technological Innovation Systems, Strategic Niche Management, or sociotechnical regimes — the frameworks through which innovation scholars routinely analyse infrastructure transitions of comparable scale and consequence. The ethnographic work that does exist is invisible to this audience, and the innovation literature has not noticed the gap (\Cref{tab:silos}).

\input{sections/tables}

This paper makes three contributions in response. First, it proposes \textit{datacentering} as a conceptual reframing — a gerund that foregrounds the active, contested, ongoing process of assembling digital infrastructure over time, rather than treating data centers as fixed technical objects amenable to efficiency optimisation. Second, it develops a multimodal research agenda structured around the conceptual and structural dimensions of these assemblages, drawing on complex adaptive systems (CAS) thinking \citep{phillips2019complex} as a sensemaking lens for understanding phenomena that are constituted by multi-actor constellations — corporations, financial investors, planning authorities, municipal governments, communities, and nonhuman actors including power grids, aquifers, and fibre networks — whose interactions produce emergent, nonlinear outcomes no single actor controls. Third, it argues that the CAS-like character of these phenomena is itself the reason why no single researcher or team can build the cartography required: the geography is too large, the data too distributed, and the opacity too deliberate. We therefore accompany this paper with an open research infrastructure — a community-maintained facility index, shared analytical pipelines, and open data — as a theoretically motivated response to the complexity of the object, and as an invitation to collective knowledge production.

The paper proceeds as follows. \Cref{sec:background} reviews the relevant literature and introduces the concept of datacentering and the CAS sensemaking lens. \Cref{sec:conceptual} develops the conceptual dimension of the research agenda, addressing how actors construct the imaginaries and territorial claims that make data center infrastructure legible or illegible. \Cref{sec:structural} addresses the structural dimension, mapping the ownership hierarchies and actor constellations — including the urban and municipal actors that the city2graph framework \citep{city2graph} makes tractable — through which governance capacity (or its absence) is constituted. \Cref{sec:temporal} addresses the temporal dimension, examining how contestation signals unfold over time and what determines whether regime actors respond. \Cref{sec:case} illustrates the framework through a comparison of Northern Virginia and Amsterdam, two markets with equivalent contestation intensity but structurally different governance responses. \Cref{sec:discussion} draws out the theoretical implications and positions the open cartography as a research programme. \Cref{sec:conclusion} closes with the invitation to build it together.
